##Introduction:
In this report, I will compare open home and open codo selling prices in San Francisco. I will use real estate data from Redfin.com, and will upload relevant Excel files onto R. I will run some statistical tests and will create visualizations for clarity.


```{r}
sfopenhomes<-read.csv("sf_open_homes.csv")
head(sfopenhomes)
```

```{r}
sfopencondos<-read.csv("sf_open_condos.csv")
head(sfopencondos)
```

```{r}
pricesqrft<-read.csv("square_foot_price.csv")
head(pricesqrft)
```

##Summary statistics of open home selling prices in San Francisco
```{r echo=TRUE}
summary(sfopenhomes$PRICE)
```
From the data, I see that the median selling price of an open home in San Francisco is 1,599,000 dollars and the mean selling price for an open home in San Francisco is 2,140,641 dollars.

##Summary statistics of open condo selling prices in San Francisco
```{r echo=TRUE}
summary(sfopencondos$PRICE)
```
From the data, I see that the median selling price of an open condo in San Francisco is 1,198,000 dollars and the mean selling price for an open condo in San Francisco is 1,398,506 dollars.

##What are the five most expensive zip codes for buying an open house in San Francisco?

```{r echo=TRUE}
library(dplyr)
sfopenhomes %>%
  group_by(ZIP) %>%
  summarise(median_price_by_zip=median(PRICE)) %>%
  arrange(desc(median_price_by_zip)) %>%
  top_n(5)

```

```{r echo=TRUE}
library(ggplot2)
df1<-data.frame(ZipCode=c("94133","94108","94115","94123","94118"),MedianPrice=c(9.5,8.75,5.145,3.924,2.89))
ggplot(df1,aes(x=ZipCode,y=MedianPrice))+geom_bar(stat="identity",fill="darkgreen")+ggtitle("Five most expensive San Francisco zip codes for open houses ")+theme(plot.title=element_text(face="bold"))+xlab("San Francisco Zip Codes")+ylab("Median home price - in millions of dollars")+geom_hline(aes(yintercept=1.599),linetype="dashed",color="red")+theme_linedraw()

```


Note-I included the median open house price of 1.599 million dollars as a dotted red line.

##What are the five least expensive zip codes for buying an open house in San Francisco?

```{r echo=TRUE}
library(dplyr)
sfopenhomes %>%
  group_by(ZIP) %>%
  summarise(median_price_by_zip=median(PRICE)) %>%
  arrange(median_price_by_zip) %>%
  top_n(-5)

```
```{r echo=TRUE}
library(ggplot2)
df2<-data.frame(ZipCode2=c("94124","94135","94112","94122","94132"),MedianPrice2=c(.9765,.988,1.049,1.2694,1.2915))
ggplot(df2,aes(x=ZipCode2,y=MedianPrice2))+geom_bar(stat="identity",fill="darkblue")+ggtitle("Five least expensive San Francisco zip codes for open houses ")+theme(plot.title=element_text(face="bold"))+xlab("San Francisco Zip Codes")+ylab("Median home price - in millions of dollars")+geom_hline(aes(yintercept=1.599),linetype="dashed",color="red")+theme_linedraw()

```


Note-I included the median open house price of 1.599 million dollars as a dotted red line.

##What are the five most expensive zip codes for buying an open condo in San Francisco?

```{r echo=TRUE}
library(dplyr)
sfopencondos %>%
  group_by(ZIP) %>%
  summarise(median_price_by_zip=median(PRICE)) %>%
  arrange(desc(median_price_by_zip)) %>%
  top_n(5)

```
```{r echo=TRUE}
library(ggplot2)
df3<-data.frame(ZipCode3=c("94108","94123","94118","94133","94114"),MedianPrice3=c(2.097,1.895,1.85,1.64,1.495))
ggplot(df3,aes(x=ZipCode3,y=MedianPrice3))+geom_bar(stat="identity",fill="darkgreen")+ggtitle("Five most expensive San Francisco zip codes for open condos ")+theme(plot.title=element_text(face="bold"))+xlab("San Francisco Zip Codes")+ylab("Median condo price - in millions of dollars")+geom_hline(aes(yintercept=1.198),linetype="dashed",color="red")+theme_linedraw()

```


-Note: Note-I included the median open condo price of 1.198 million dollars as a dotted red line.

##What are the five least expensive zip codes for buying an open condo in San Francisco?

```{r echo=TRUE}
library(dplyr)
sfopencondos %>%
  group_by(ZIP) %>%
  summarise(median_price_by_zip=median(PRICE)) %>%
  arrange(median_price_by_zip) %>%
  top_n(-5)

```
```{r echo=TRUE}
library(ggplot2)
df4<-data.frame(ZipCode4=c("94132","94111","94134","94124","94116"),MedianPrice4=c(.6485,.699,.785,.799,.949))
ggplot(df4,aes(x=ZipCode4,y=MedianPrice4))+geom_bar(stat="identity",fill="darkblue")+ggtitle("Five least expensive San Francisco zip codes for open condos ")+theme(plot.title=element_text(face="bold"))+xlab("San Francisco Zip Codes")+ylab("Median open condo price - in millions of dollars")+geom_hline(aes(yintercept=1.198),linetype="dashed",color="red")+theme_linedraw()

```


-Note: Note-I included the median open condo price of 1.198 million dollars as a dotted red line.

##Running a statistical test
I now will conduct a t-test to determine if there is a significant difference between the mean price per square foot for open homes and that of condos
```{r echo=TRUE}
t.test(pricesqrft$PSQFT~pricesqrft$Property_Type)
```

After running the two-sample t-test, I observe that the mean price per square foot for an open condo is significantly more than that for an open home. On average, open condos cost more per square foot than open homes in San Francisco.

##Visualization of pricing distributions (per square foot) for condos and homes.

```{r echo=TRUE}
library(ggplot2)
ggplot(pricesqrft, aes(x=PSQFT))+geom_histogram(fill="red")+facet_wrap(~Property_Type)+theme_linedraw()
```
```{r echo=TRUE}
library(ggplot2)
ggplot(pricesqrft,aes(x=PSQFT, fill=Property_Type))+geom_density(alpha=0.5)+theme_linedraw()
```


##Conclusion
From my analysis, despite the fact that open homes cost more than open condos, I see how open houses are actually better deals than open condos in the San Francisco real estate market.
